---
title: "Measles serology"
number-sections: true
format:
  html:
    toc: true
    toc-depth: 4
editor: source
editor_options: 
  chunk_output_type: console
#bibliography: references.bib
#csl: the-american-naturalist.csl
---

```{r include = FALSE, eval = FALSE}
# The code that adds the .nojekyll file to the GitHub repository:
file <- ".nojekyll"
file.create(file)
gert::git_add(file)
gert::git_commit("Adding the .nojekyll file")
gert::git_push()
rm(file)
```

```{r include = FALSE, eval = FALSE}
run_all <- function() {
  knitr::purl("index.qmd", "tmp.R", documentation = FALSE)
  source("tmp.R")
  file.remove("tmp.R")  
}

rerun_all <- function() {
  rm(list  = grep("run_all", ls(envir = .GlobalEnv), value = TRUE, invert = TRUE),
     envir = .GlobalEnv)
  run_all()
}
```

```{r include = FALSE, purl = FALSE}
setHook("plot.new", NULL, action = "replace")

par3 <- function(...) par(..., mgp = c(1.5, .5, 0), bty = "n")

knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par3(plt = c(.105, .97, .15, .95)) else NULL
  })

figw <- 7  # default is 7 and maximum possible is 8.3
figh <- 5  # default is 5
eps <- .8

knitr::opts_chunk$set(margin     = TRUE,
                      fig.retina = 2,
                      fig.align  = "center",
                      fig.width  = eps * figw,
                      fig.height = eps * figh) 
```


## Constants

The path to the data folder:

```{r}
if (grepl("arc|hoisy", Sys.getenv("USER"))) {
  path2data <- paste0("~/Library/CloudStorage/OneDrive-OxfordUniversityClinical",
                      "ResearchUnit/GitHub/choisy/measles serology/")
}
```

```{r include = FALSE}
path2cache <- paste0(path2data, "cache/")
if (! dir.exists(path2cache)) dir.create(path2cache)
make_path <- function(x) paste0(path2cache, x)
file_exists <- function(x) file.exists(make_path(x))
readRDScache <- function(x) readRDS(make_path(x))
saveRDScache <- function(object, file) saveRDS(object, make_path(file))
```


## Packages

Required packages:

```{r}
required <- c("sf", "readxl", "stringr", "purrr", "tidyr", "magrittr", "dplyr",
              "lubridate", "mgcv")
```

Installing those that are not installed yet:

```{r}
to_install <- required[! required %in% installed.packages()[, "Package"]]
if (length(to_install)) install.packages(to_install)
```

Loading the packages for interactive use:

```{r warning = FALSE, message = FALSE}
invisible(sapply(required, library, character.only = TRUE))
```


## Functions

A function that generates names of a vector of character strings from its content:

```{r}
set_names2 <- function(x) {
  setNames(x, x |>
                str_remove("^.*_") |> 
                str_remove("^.*-") |> 
                str_remove(".xlsx$"))
}
```

A tuning of the `text()` function that includes white background to the text:

```{r}
text2 <- function(x, y, labels, cex = 1, bg = "white", padx = 10, pady = .1, ...) {
  w <- strwidth(labels, cex = cex)
  h <- strheight(labels, cex = cex)
  rect(x - w / 2 - padx, y - h / 2 - pady / 2, x + w / 2 + padx, y + h / 2 + pady / 2,
       col = bg, border = NA)
  text(x, y, labels, cex = cex, ...)
}
```

Tuning of the `rgb()` function:

```{r}
rgb2 <- function(...) rgb(..., maxColorValue = 255)
```

Tunings of the `plot()` function:

```{r}
plot2 <- function(...) plot(..., axes = FALSE)
plot3 <- function(...) plot2(..., ann = FALSE)
plot4 <- function(...) plot3(..., type = "n")
```

A tuning of the `pie()` function:

```{r}
pie2 <- function(x, ...) if (all(x == 0)) plot4(1) else pie(x, ...)
```

A function that pads a data frame of coordinates:

```{r}
pad <- function(x) {
  x <- arrange(x, 1)
  xs <- x[[1]]
  tibble(c(first(xs), xs, last(xs)), c(0, x[[2]], 0)) |> 
    setNames(names(x)[1:2])
}
```

Some tunings of the `abline()` function:

```{r}
abline2 <- function(...) abline(..., col = "lightgrey")
abline3 <- function(...) abline(..., col = "grey")
```

Tuning of the `grep()` function:

```{r}
grep2 <- function(pattern, x, ...) grep(x, pattern, value = TRUE, invert = TRUE, ...)
```

A tuning of the `mgcv::gam()` function:

```{r}
gam2 <- function(...) mgcv::gam(..., method = "REML")
```

A function that retrieves the REML from a GAM model:

```{r}
get_reml <- function(x) as.numeric(x$gcv.ubre)
```

A function that retrieves the percentage of deviance explained from a GAM model:

```{r}
dev_explained <- function(x) summary(x)$dev.expl
```

A tuning of the `segments()` function:

```{r}
segments2 <- function(...) segments(..., col = 2, lwd = 4)
```

A function that extends the `tally()` function by adding a column of the corresponding
proportions:

```{r}
tally2 <- function(x) {
  x |> 
    tally() |> 
    mutate(prop = 100 * n / sum(n))
}
```

A tuning of the `arrows()` function:

```{r}
arrows2 <- function(...) arrows(..., angle = 90, code = 3)
```

A tuning of the `points()` function:

```{r}
points2 <- function(...) points(..., pch = 19)
```

A tuning of the `par()` function:

```{r}
par2 <- function(...) par(..., cex = 1)
```


## Data

### Functions

A function that recodes sex:

```{r}
recode_sex <- function(x) {
  sex_val <- sort(unique(x$sex))
  males   <- grep("^M|na", sex_val, TRUE, value = TRUE)
  females <- grep("^M|na", sex_val, TRUE, value = TRUE, invert = TRUE)
  mutate(x, across(sex, ~ recode_values(.x, males ~ "male", females ~ "female")))
}
```

A piece of code common to all the sample collections:

```{r}
common_data_processing <- function(x) {
  x |> 
    filter_out(if_all(everything(), is.na)) |>
    recode_sex() |>
    separate_wider_delim(ID, regex("\\s+"), names = c("sID", "cdate", "plate")) |>
    mutate(across(titer, as.double))
}
```

### Data

Loading all the excel files of the data folder:

```{r}
the3files <- path2data |> 
  list.files(full.names = TRUE) |> 
  grep2("gadm") |> 
  grep2("cache") |> 
  set_names2() |> 
  map(read_excel, "Marc", col_types = "text")
```

The first collection:

```{r}
first_collection <- the3files$DEC2019 |> 
  common_data_processing() |> 
  mutate(date_chr = if_else( grepl("[./]", date_collection), date_collection, NA),
         date_num = if_else(!grepl("[./]", date_collection), date_collection, NA),
         across(date_num, ~ .x |> as.numeric() |> as_date("1899-12-30")),
         across(date_chr, dmy),
         collection_date = coalesce(date_num, date_chr)) |> 
  select(participant, hospital, collection, sample_ID, age_group, age, sex,
         collection_date, plate, titer, conclusion)
```

Combining the collections:

```{r warning = FALSE}
dataset <- the3files |> 
  extract(c("Dec2022", "Apr24")) |> 
  map(select, -date_collection) |> 
  bind_rows() |> 
  common_data_processing() |> 
  mutate(collection_date = ymd(paste(year_collection, month_collection, day_collection,
                                     sep = "-"))) |> 
  select(participant, province, hospital, site, collection, sample_ID, age_group, age,
         sex, collection_date, plate, titer, conclusion) |> 
  bind_rows(first_collection) |> 
  mutate(province = ifelse(is.na(province), hospital, province),
         across(province, ~ recode_values(.x, "AG"     ~ "An Giang",
                                              "BD"     ~ "Binh Dinh",
                                              "DL"     ~ "Dak Lak",
                                              "DT"     ~ "Dong Thap",
                                              "HC"     ~ "HCMC",
                                              "HU"     ~ "Hue",
                                              "KG"     ~ "Kien Giang",
                                              "KH"     ~ "Khanh Hoa",
                                              "QN"     ~ "Quang Ngai",
                                              "ST"     ~ "Soc Trang",
                                              "Daklak" ~ "Dak Lak",
                                              "HUE"    ~ "Hue",
                                              "HCM"    ~ "HCMC", default = .x)),
         year = year(collection_date),
         time_point = recode_values(year, 2019:2020 ~ "T1",
                                          2022:2023 ~ "T2", default = "T3"),
         across(age, ~ ifelse(participant == "0158" & province == "Dak Lak" &
                                collection == "42", "48", .x)),
         age_unit = str_remove_all(age, "\\d"),
         age2 = as.numeric(str_remove_all(age, "\\D")),
         age3 = as.numeric(age),  # note: this generate NAs
         age4 = ifelse(is.na(age3), ifelse(age_unit %in% c("M", "th"), age2 / 12,
                                           ifelse(age_unit == "ng", age2 / 365, age2)),
                       age3),
         age = ifelse(age4 > 100, year - age4, age4),
         age_group = cut(age, c(0:20, seq(25, 70, 5), 100), right = FALSE),
         across(conclusion, ~ ifelse(.x == "pós", "Pos", .x))) |> 
  select(-c(year, age_unit, age2, age3, age4, hospital)) |> 
  arrange(time_point, province, age_group, collection_date)
```

A hash table to generate site ID from province names:

```{r}
hash <- dataset |> 
  filter_out(time_point == "T1") |> 
  filter(age > 20) |> 
  select(province, site) |> 
  unique() |> 
  with(setNames(site, province))
```

Completing the site ID when missing:

```{r}
dataset <- mutate(dataset, across(site, ~ ifelse(is.na(.x), hash[province], .x)))
```

Which gives:

```{r}
dataset
```

Checking the thresholds:

```{r}
dataset |> 
  filter(conclusion == "Neg") |> 
  pull(titer) |> 
  max()

dataset |> 
  filter(conclusion == "Equi") |> 
  pull(titer) |> 
  range()

dataset |> 
  filter(conclusion == "Pos") |> 
  pull(titer) |> 
  min()
```


## Sampling locations

Loading the `sf` object of the provinces of Vietnam:

```{r message = FALSE}
gadm1 <- readRDS(paste0(path2data, "gadm36_VNM_1_sf.rds"))
st_crs(gadm1) <- st_crs(gadm1)
```

Plotting the provinces from which samples where collected:

```{r fig-map, fig.width = .7 * 7, fig.height = 6}
sel <- dataset |>
  pull(province) |> 
  unique() |> 
  str_replace("Hue", "Thua Thien Hue") |> 
  str_replace("HCMC", "Ho Chi Minh")

opar <- par(plt = rep(0:1, 2))

gadm1 |> 
  st_geometry() |> 
  plot(col = "lightyellow")

gadm1 |> 
  filter(VARNAME_1 %in% setdiff(sel, "Kien Giang")) |> 
  st_geometry() |> 
  plot(add = TRUE, col = 2)

gadm1 |> 
  filter(VARNAME_1 == "Kien Giang") |> 
  st_geometry() |> 
  plot(add = TRUE, col = 7)

par(opar)
```


## Samples metadata

Number of samples per site per time point:

```{r eval = FALSE}
with(dataset, table(province, time_point))
```

```{r echo = FALSE}
(tmp <- with(dataset, table(province, time_point)))
saveRDScache(tmp, "nb_samples_prov_timepoint.rds")
```

A tuning of the `barplot()` function:

```{r}
make_barplot2 <- function(n1 = 5, n2 = 10, ad = 20) {
  function(x, ...) {
    with(x, {
      barplot(Freq, space = 0, col = 4, ...)
      mtext(province[1], line = -.5)
    })
    segments2( 0, n1, ad, n1)
    segments2(ad, n2, 31, n2)
  }
}

barplot2 <- make_barplot2()
```

Number of sample per age class, per province, per time point:

```{r fig-nb_samples_site_timepoint_ageclass, fig.width = 7, fig.height = 50 / 3}
tmp <- dataset |> 
  with(table(age_group, province, time_point)) |>
  as.data.frame()

ylim <- c(0, max(tmp$Freq))

opar <- par2(mfcol = c(10, 3), plt = c(.12, .985, .1, .9))
tmp |> 
  group_by(time_point, province) |> 
  group_split() |> 
  walk(~ barplot2(.x, ylim = ylim))
par(opar)
```

Number of samples per day and site:

```{r fig-nb_sample_day_site}
hz_lines <- 1:10

opar <- par(plt = c(.105, .97, .1, .9))

dataset |> 
  group_by(province, collection_date) |> 
  tally() |> 
  ungroup() |> 
  mutate(y = max(hz_lines) + 1 - as.numeric(as.factor(province))) |> 
  with(plot3(collection_date, y, cex = sqrt(n), col = 4, xpd = TRUE))

lb <- 2020:2024
axis(1, ymd(paste(lb, "0101")), lb)
abline3(h = hz_lines)
text2(rep(ymd(20210701), length(hz_lines)), rev(hz_lines), unique(dataset$province))

par(opar)
```

An alternative representation:

```{r fig-cum_nb_sample}
plot2(NA, xlim = range(dataset$collection_date), ylim = c(0, 600), xlab = NA,
      ylab = "cumulative number of samples")
axis(1, ymd(paste(lb, "0101")), lb); axis(2)

dataset |> 
  group_by(province, collection_date) |> 
  tally() |> 
  group_modify(~ mutate(.x, cumn = cumsum(n))) |> 
  group_walk(~ with(.x, lines(collection_date, cumn, type = "s", col = 4)))
```

Yet another representation:

```{r fig-sampling_intensity}
dataset |> 
  group_by(collection_date) |> 
  tally() |> 
  with(plot(collection_date, n, type = "h", col = "4", xlab = NA,
            ylab = "sampling intensity (/day)"))
```

Gender distribution:

```{r fig-sex_distribution, fig.width = .7 * 7, fig.height = .7 * 50 / 3}
opar <- par(mfcol = c(10, 3), plt = c(.05, .95, .01, .91))
dataset |> 
  with(table(sex, province, time_point)) |> 
  as.data.frame() |> 
  group_by(time_point, province) |> 
  group_split() |> 
  walk(~ with(.x, {
    pie2(Freq, NA, col = c(rgb2(247, 114, 184), rgb2(62, 146, 247)))
    mtext(province[1], line = -.5)
  }))
par(opar)
```


## Titer distribution

A function that plots a 3-color distribution from a 2-column data frame `df` that
contains the coordinates of the distribution, with $x$-values in the first column
named `x` and $y$-values in the second column, on a decimal log scale of the $x$
values:

```{r}
ptiters <- function(df, main = NULL, th1 = 150, th2 = 200, lb2 = 1:5, ...) {
  plot2(df, type = "n", xlab = "titer (IU/L)", ylab = "density", ...)
  lb1 <- unlist(map(10^(lb2), ~ .x * 1:10))
  axis(1, log10(lb1), NA, tcl = - .25); axis(1, lb2, 10^lb2); axis(2)
  df |> 
    group_by(cut(x, c(-Inf, log10(c(th1, th2)), Inf))) |> 
    group_map(~ pad(.x)) |> 
    walk2(c(2, 7, 3), ~ polygon(.x, col = .y, border = NA))
  if (! is.null(main)) title(main, line = -.25, cex.main = 1)
}
```

A function that (1) generates the 2-column input data frame of the function
`ptiters()` from the column named `titer` of the data frame `df` and then (2) feeds it
to the `ptiters()` function to draw the 3-color distribution:

```{r}
plot_titers <- function(df, main = NULL, ...) {
  df |> 
    pull(titer) |> 
    log10() |> 
    density() |> 
    extract(1:2) |> 
    as_tibble() |> 
    ptiters(main, ...)
}
```

Plotting the distribution of all the titers, across the 3 time points and the 10 sites:

```{r fig-distribution_titers}
plot_titers(dataset)
```

The corresponding proportions in each of the 3 categories of titer values are the
following:

```{r}
dataset |> 
  group_by(conclusion) |> 
  tally2()
```

The distributions of all the titers across the 10 sites for each time point separately:

```{r fig-distribution_titers_timepoints, fig.width = .7 * eps * figw, fig.height = .7 * 3 * eps * figh}
opar <- par2(mfrow = c(3, 1), plt = c(.13, .97, .18, .95))
dataset |> 
  group_by(time_point) |> 
  group_split() |> 
  walk2(c(paste("Dec", c(2019, 2022)), "Apr 2024"),
        ~ plot_titers(.x, .y, xlim = c(.5, 4.5), ylim = c(0, .8)))
par(opar)
```

The corresponding proportions in each of the 3 categories of titer values and for each
time point are the following:

```{r}
dataset |> 
  group_by(time_point, conclusion) |> 
  tally2() |> 
  pivot_wider(id_cols = -n, names_from = time_point, values_from = prop)
```

### Additional

```{r fig-titers_raw_data}
with(dataset, plot(age, log10(titer), col = adjustcolor("black", .2)))
```


## Sero-prevalence

This function takes a data frame `x` as input that contains at least 2 columns named
`conclusion` and `n` that are character and numeric values respectively. The
`conclusion` column as one of its values equal to `Pos`. The output of the function is
a named vector of 3 values being the estimate of the proportion of `Pos` and the lower
and upper bounds of the 95% confidence interval of this estimate.

```{r}
estimate_prop <- function(x) {
  slot_names <- c("p", "l", "u")
  x <- with(x, setNames(n, conclusion))
  n <- sum(x)
  if (n < 1) return(setNames(rep(NA, 3), slot_names))
  binom.test(x["Pos"], n) |> 
    extract(c("estimate", "conf.int")) |> 
    unlist() |> 
    setNames(slot_names)
}
```

This function takes a data frame that contains at least 2 columns named `age_group`, 
`conclusion` and as many rows as observations. It returns a data frame with 3 columns
(proportion estimate and confidence interval bounds) and as many rows as age groups:

```{r}
make_prepare_data <- function(template = dataset |> 
                                group_by(age_group, conclusion) |> 
                                tally() |> 
                                ungroup() |> 
                                select(-n)) {
  function(x) {
    x |> 
      group_by(age_group, conclusion) |> 
      tally() |> 
      left_join(x = template, y = _, c("age_group", "conclusion")) |> 
      replace_na(list(n = 0)) |> 
      group_by(age_group) |> 
      group_split() |> 
      map_dfr(estimate_prop)
  }
}

prepare_data <- make_prepare_data()
```

This function takes the ouptut of `prepare_data()` and plots the sero-prevalence per
age group:

```{r}
make_p_age_profile <- function(xs = c(.5:19.5, seq(22.5, 67.5, 5), 75)) {
  function(x) {
    plot(xs, rep(.5, length(xs)), ylim = 0:1, type = "n", xaxs = "i", yaxs = "i",
         xlim = c(0, 80), xlab = "age classe (years)", ylab = "sero-prevalence")
    abline2(v = seq(0, 70, 5))
    abline2(h = seq(0, 1, .1))
    with(x, {
      arrows2(xs, l, xs, u, .02, col = 4)
      points2(xs, p, col = 4)
    })
    box(bty = "o")
  }
}

p_age_profile <- make_p_age_profile()
```

Wrapping up `prepare_data()` and `p_age_profile()`:

```{r}
plot_age_profile <- function(x) {
  x |> 
    prepare_data() |> 
    p_age_profile()
}
```

The age profile of sero-prevalence across the 3 time points and the 10 sites:

```{r fig-age_profile}
plot_age_profile(dataset)
```

The age profile of sero-prevalence across the 10 sites and for the 3 time points
separately:

```{r fig-age_profile_timepoints, fig.width = .7 * eps * figw, fig.height = .7 * 3 * eps * figh}
opar <- par2(mfrow = c(3, 1), plt = c(.13, .97, .18, .95))
dataset |> 
  group_by(time_point) |> 
  group_walk(~ plot_age_profile(.x))
par(opar)
```

`x` is a list of data frames of 3 columns (proportion estimates and confidence interval
bounds). The number of rows of the data frames is equal to the number age group and the
number of slot of the list `x` is equal to the number of time points. The following
function returns a list of data frames of 3 columns too where the number of slots of
the list and the number of rows of the data frames are switched.

```{r}
time_change <- function(x) {
  map(1:nrow(first(x)), function(y) map_dfr(x, ~ .x[y, ]))
}
```

This function takes a 3-column (proportion estimate and confidence interval bounds) /
3-row (for 3 time points) data frame that corresponds to a given age group and plots 
the change in sero-prevalence as a function of time.

```{r}
time_trend <- function(x, xs = 1:3, col = 4) {
  with(x, {
    plot4(xs, p, xlim = c(0, 4), ylim = 0:1)
    abline2(h = seq(.2, .8, .2))
    arrows2(xs, l, xs, u, .05, col = col)
    points2(xs, p, col = col)
    box(bty = "o")
  })
}
```

Showing the change in seroprevalence as a function of time for all the age groups
separately:

```{r fig-seroprevalence_trends, fig.width = 8.4, fig.height = 1.8}
epsl <- .065
opar <- par(mfrow = c(3, 10), plt = rep(c(epsl, 1 - epsl), 2))
dataset |> 
  filter_out(age_group == last(levels(age_group))) |> 
  group_by(time_point) |> 
  group_split() |> 
  map(prepare_data) |> 
  map(~ .x[-nrow(.x), ]) |> 
  time_change() |> 
  walk(time_trend)
par(opar)
```

### Additional

```{r}
data2 <- mutate(dataset, log_titer = log10(titer),
                positive1 = titer > 200,
                positive2 = titer > 150)
```

#### Functions

```{r}
predict2 <- function(model, x = seq(0, 70, le = 512)) {
  bind_cols(tibble(x  = x),
            model |> 
              predict(list(age = x), se.fit = .95) |> 
              map(model$family$linkinv) |> 
              as_tibble())
}
```

```{r}
add_ribbon <- function(pred, col = 4) {
  with(pred, {
    polygon(c(x, rev(x)), c(ll, rev(ul)), border = NA, col = adjustcolor(col, .2))
    lines(x, fit, col = col, lwd = 2)
  })
}
```

```{r}
plot_pred <- function(pred, ylog = TRUE, ylab = "mean titer (IU/L)", ylim = NULL,
                      col = 4, ...) {
  with(pred, {
    if (is.null(ylim)) ylim <- pred |> 
        select(-x) |> 
        range()
    plot2(x, fit, type = "n", ylim = ylim, xaxs = "i", ylab = ylab,
          xlab = "age (years)", ...)
  })
  if (ylog) {
    lb1 <- unlist(map(10^(1:5), ~ .x * 1:10))
    lb2 <- c(150, 200, 300, 500, 700, 1000)
    axis(2, log10(lb1), NA)
    axis(2, log10(lb2), lb2)
  } else {
    axis(2)
  }
  axis(1)
  abline2(v = c(.75, 1.5, seq(5, 65, 10)))
  abline3(v = seq(10, 60, 10))
  add_ribbon(pred, col)
  box(bty = "o")
}
```

```{r}
plot_seroprevalences <- function(data) {
  gam2(positive1 ~ s(age, bs = "tp", k = 50), binomial, data) |> 
    predict2() |> 
    plot_pred(ylog = FALSE, ylab = "sero-prevalence", ylim = 0:1, yaxs = "i", col = 3)
  
  gam2(positive2 ~ s(age, bs = "tp", k = 50), binomial, data) |> 
    predict2() |> 
    add_ribbon(2)
  
  abline3(h = 1:9 / 10)
}
```


#### Analysis

```{r eval = FALSE}
models_prev <- tibble(
  smoothing_basis = c("thin plate", "thin plate", "Duchon", "cubic regression",
                      "cubic regression", "cyclic cubic", "B-spline", "P-spline",
                      "cyclic P-spline", "Gaussian process"),
  code = c("tp", "ts", "ds", "cr", "cs", "cc", "bs", "ps", "cp", "gp")) |> 
  mutate(model   = map(code, ~ gam2(positive1 ~ s(age, bs = .x, k = 50),
                                    binomial, data2)),
         REML    = map_dbl(model, get_reml),
         per_dev = map_dbl(model, dev_explained),
         AIC     = map_dbl(model, AIC),
         k_check = map(model, ~ as_tibble(k.check(.x, n.rep = 1000)))) |> 
  unnest_wider(k_check)
```

```{r include = FALSE}
if (file_exists("models_prev.rds")) {
  models_prev <- readRDScache("models_prev.rds")
} else {
  models_prev <- tibble(
    smoothing_basis = c("thin plate", "thin plate", "Duchon", "cubic regression",
                        "cubic regression", "cyclic cubic", "B-spline", "P-spline",
                        "cyclic P-spline", "Gaussian process"),
    code = c("tp", "ts", "ds", "cr", "cs", "cc", "bs", "ps", "cp", "gp")) |> 
    mutate(model   = map(code, ~ gam2(positive1 ~ s(age, bs = .x, k = 50),
                                      binomial, data2)),
           REML    = map_dbl(model, get_reml),
           per_dev = map_dbl(model, dev_explained),
           AIC     = map_dbl(model, AIC),
           k_check = map(model, ~ as_tibble(k.check(.x, n.rep = 1000)))) |> 
    unnest_wider(k_check)
  saveRDScache(models_prev, "models_prev.rds")
}
```

which gives:

```{r}
select(models_prev, -model)
```

Let's then stick to the thin plate smoothing. Plotting smoothed trends:

```{r fig-gam_seroprev}
models_prev |> 
  filter(code == "tp") |> 
  pull(model) |> 
  first() |>
  predict2() |>
  plot_pred(ylog = FALSE, ylab = "prevalence", ylim = 0:1, yaxs = "i")
abline3(h = 1:9 / 10)
```

```{r fig-gam_seroprev2}
plot_seroprevalences(data2)
```


## GAM models

All the 10 smoothing bases more or less perform equally well:

```{r include=FALSE, eval = FALSE}
models <- tibble(
  smoothing_basis = c("thin plate", "thin plate", "Duchon", "cubic regression",
                      "cubic regression", "cyclic cubic", "B-spline", "P-spline",
                      "cyclic P-spline", "Gaussian process"),
  code = c("tp", "ts", "ds", "cr", "cs", "cc", "bs", "ps", "cp", "gp")) |> 
  mutate(model   = map(code, ~ gam2(log_titer ~ s(age, bs = .x, k = 50),
                                    data = data2)),
         REML    = map_dbl(model, get_reml),
         per_dev = map_dbl(model, dev_explained),
         AIC     = map_dbl(model, AIC),
         k_check = map(model, ~ as_tibble(k.check(.x, n.rep = 1000)))) |> 
  unnest_wider(k_check)
```

```{r include = FALSE, eval=FALSE}
if (file_exists("models.rds")) {
  models <- readRDScache("models.rds")
} else {
  models <- tibble(
    smoothing_basis = c("thin plate", "thin plate", "Duchon", "cubic regression",
                        "cubic regression", "cyclic cubic", "B-spline", "P-spline",
                        "cyclic P-spline", "Gaussian process"),
    code = c("tp", "ts", "ds", "cr", "cs", "cc", "bs", "ps", "cp", "gp")) |> 
    mutate(model   = map(code, ~ gam2(log_titer ~ s(age, bs = .x, k = 50),
                                      data = data2)),
           REML    = map_dbl(model, get_reml),
           per_dev = map_dbl(model, dev_explained),
           AIC     = map_dbl(model, AIC),
           k_check = map(model, ~ as_tibble(k.check(.x, n.rep = 1000)))) |> 
    unnest_wider(k_check)
  saveRDScache(models, "models.rds")
}
```

Which gives:

```{r include=FALSE, eval=FALSE}
select(models, -model)
```

Let's then stick to the thin plate smoothing. Plotting smoothed trends:

```{r fig-gam_titer, include=FALSE, eval=FALSE}
models |> 
  filter(code == "tp") |> 
  pull(model) |> 
  first() |>
  predict2() |> 
  plot_pred()
abline(h = log10(c(150, 200)), col = 2:3, lty = 2, lwd = 2)
```



## TO DO

* model density() in 2D in order to see density of data points not only as a function 
of age but also as a function of titers

```{r}
image_titer_density_age <- function(data, xaxis = TRUE, xlab = "age (years)",
                                    min_titer = NULL, max_titer = NULL) {
  if (is.null(min_titer)) min_titer <- min(data$log_titer)
  if (is.null(max_titer)) max_titer <- max(data$log_titer)
  
  data |> 
    filter(age < 70) |> 
    group_by(age_group) |> 
    group_map(~ density(.x$log_titer, from = min_titer, to = max_titer)$y) |> 
    rep(c(rep(1, 20), rep(5, 10))) |> 
    as.data.frame() |> 
    as.matrix() |> 
    t() |>
    image(1:70 - .5, seq(min_titer, max_titer, le = 512), z = _, axes = FALSE,
          xlab = xlab, ylab = "titer (IU/L)",
          col = hcl.colors(256, "YlOrRd", rev = TRUE))
  
  if (xaxis) axis(1)
  lb1 <- unlist(map(10^(1:5), ~ .x * 1:10))
  lb2 <- 1:5
  axis(2, log10(lb1), NA, tcl =  - .25)
  axis(2, lb2, 10^lb2)
  
  abline(h = log10(c(150, 200)), col = 2:3, lty = 2, lwd = 2)
  
  box(bty = "o")
}
```

```{r fig-titer_density_age}
image_titer_density_age(data2)
```

```{r fig-titers_seroprevalences, fig.height = 2 * eps * figh}
opar <- par(plt = c(.105, .97, .5, 1 - (1 - .95) / 2 - .15 / 2))
image_titer_density_age(data2, xaxis = FALSE, xlab = NA, min_titer = log10(5))
axis(3, 10 * 0:7)
mtext("age (years)", line = 1.7)
par(plt = c(.105, .97, .15 / 2 + (1 - .95) / 2, .5), new = TRUE)
plot_seroprevalences(data2)
par(opar)
```


* doing clustering on the ANCOVA

```{r fig-T1T2T3}
data2 |> 
  filter(time_point == "T1") |> 
  gam2(positive1 ~ s(age, bs = "tp", k = 50), binomial, data = _) |> 
  predict2() |> 
  plot_pred(ylog = FALSE, ylab = "sero-prevalence", ylim = 0:1, yaxs = "i", col = 2)

data2 |> 
  filter(time_point == "T2") |>
  gam2(positive1 ~ s(age, bs = "tp", k = 50), binomial, data = _) |> 
  predict2() |> 
  add_ribbon(3)

data2 |> 
  filter(time_point == "T3") |>
  gam2(positive1 ~ s(age, bs = "tp", k = 50), binomial, data = _) |> 
  predict2() |> 
  add_ribbon(4)
```

```{r fig-T1T2}
data2 |> 
  filter(time_point == "T1") |> 
  gam2(positive1 ~ s(age, bs = "tp", k = 50), binomial, data = _) |> 
  predict2() |> 
  plot_pred(ylog = FALSE, ylab = "sero-prevalence", ylim = 0:1, yaxs = "i", col = 2)

data2 |> 
  filter(time_point == "T2") |>
  gam2(positive1 ~ s(age, bs = "tp", k = 50), binomial, data = _) |> 
  predict2() |> 
  add_ribbon(3)
```

```{r}
seroprev3timepoints <- function(x, k = 50, title = FALSE) {
  x |> 
    filter(time_point == "T1") |> 
    gam2(positive1 ~ s(age, bs = "tp", k = k), binomial, data = _) |> 
    predict2() |> 
    plot_pred(ylog = FALSE, ylab = "sero-prevalence", ylim = 0:1, yaxs = "i", col = 2)
  
  x |> 
    filter(time_point == "T2") |>
    gam2(positive1 ~ s(age, bs = "tp", k = k), binomial, data = _) |> 
    predict2() |> 
    add_ribbon(3)
  
  x |> 
    filter(time_point == "T3") |>
    gam2(positive1 ~ s(age, bs = "tp", k = k), binomial, data = _) |> 
    predict2() |> 
    add_ribbon(4)
  
  if (title) x |> 
    pull(province) |> 
    first() |> 
    text(35, .2, labels = _)
}
```

```{r fig-seroprev_provinces_3timepoints, fig.width = 8.4, fig.height = 5 * 8.4 / 7}
opar <- par(mfrow = c(3,3), plt = c(.12, .97, .18, .95))
data2 |> 
  filter_out(province == "Kien Giang") |> 
  group_by(province) |> 
  group_split() |> 
  walk(seroprev3timepoints, title = TRUE)
par(opar)
```

```{r fig-seroprev_provinces, fig.width = 8.4, fig.height = 5 * 8.4 / 7}
opar <- par(mfrow = c(3,3), plt = c(.12, .97, .18, .95))
dataset |> 
  filter_out(province == "Kien Giang") |> 
  group_by(province) |> 
  group_split() |> 
  walk(plot_age_profile)
par(opar)
```


* understanding the age profile of titers and seroprevalence. Maybe as an effect of
vaccination and waning immunity for vaccination

* Looking at 0-1 and 0-5 in order to see whether we see more detail signal as a
function of age


