---
title: "Measles serology"
number-sections: true
format:
  html:
    toc: true
    toc-depth: 4
editor: source
editor_options: 
  chunk_output_type: console
#bibliography: references.bib
#csl: the-american-naturalist.csl
---

```{r include = FALSE, eval = FALSE}
# The code that adds the .nojekyll file to the GitHub repository:
file <- ".nojekyll"
file.create(file)
gert::git_add(file)
gert::git_commit("Adding the .nojekyll file")
gert::git_push()
rm(file)
```

```{r include = FALSE, eval = FALSE}
run_all <- function() {
  knitr::purl("index.qmd", "tmp.R", documentation = FALSE)
  source("tmp.R")
  file.remove("tmp.R")  
}

rerun_all <- function() {
  rm(list  = grep("run_all", ls(envir = .GlobalEnv), value = TRUE, invert = TRUE),
     envir = .GlobalEnv)
  run_all()
}
```

```{r include = FALSE, purl = FALSE}
setHook("plot.new", NULL, action = "replace")

par2 <- function(...) par(..., mgp = c(1.5, .5, 0), bty = "n")

knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par2(plt = c(.105, .97, .15, .95)) else NULL
  })

figw <- 7  # default is 7 and maximum possible is 8.3
figh <- 5  # default is 5
eps <- .7

knitr::opts_chunk$set(margin     = TRUE,
                      fig.retina = 2,
                      fig.align  = "center",
                      fig.width  = eps * figw,
                      fig.height = eps * figh) 
```


## Contants

The path to the data folder:

```{r}
if (grepl("arc|hoisy", Sys.getenv("USER"))) {
  path2data <- paste0("~/Library/CloudStorage/OneDrive-OxfordUniversityClinical",
                      "ResearchUnit/GitHub/choisy/measles serology/")
}
```

```{r include = FALSE}
path2cache <- paste0(path2data, "cache/")
if (! dir.exists(path2cache)) dir.create(path2cache)
make_path <- function(x) paste0(path2cache, x)
saveRDScache <- function(object, file) saveRDS(object, make_path(file))
```


## Packages

Required packages:

```{r}
required <- c("sf", "readxl", "stringr", "purrr", "tidyr", "magrittr", "dplyr",
              "lubridate")
```

Installing those that are not installed yet:

```{r}
to_install <- required[! required %in% installed.packages()[,"Package"]]
if (length(to_install)) install.packages(to_install)
```

Loading the packages for interactive use:

```{r warning = FALSE, message = FALSE}
invisible(sapply(required, library, character.only = TRUE))
```


## Functions

A function that generates names of a vector of character strings from its content:

```{r}
set_names2 <- function(x) {
  setNames(x, x |>
             str_remove("^.*_") |> 
             str_remove("^.*-") |> 
             str_remove(".xlsx$"))
}
```

A tuning of the `text()` function:

```{r}
text2 <- function(x, y, labels, cex = 1,
                          bg = "white", padx = 10, pady = .1, ...) {
  w <- strwidth(labels, cex = cex)
  h <- strheight(labels, cex = cex)
  rect(x - w / 2 - padx, y - h / 2 - pady / 2, x + w / 2 + padx, y + h / 2 + pady / 2,
       col = bg, border = NA)
  text(x, y, labels, cex = cex, ...)
}
```

Tuning of the `rgb()` function:

```{r}
rgb2 <- function(...) rgb(..., maxColorValue = 255)
```

A tuning of the `pie()` function:

```{r}
pie2 <- function(x, ...) {
  if (all(x == 0)) {
    plot(1, ann = FALSE, axes = FALSE, type = "n")
  } else {
    pie(x, ...)
  }
}
```

A function that pads a data frame of coordinates:

```{r}
pad <- function(x) {
  x <- arrange(x, 1)
  xs <- x[[1]]
  tibble(c(first(xs), xs, last(xs)), c(0, x[[2]], 0)) |> 
    setNames(names(x)[1:2])
}
```

Tuning of the `abline()` function:

```{r}
abline2 <- function(...) abline(..., col = "lightgrey")
```


## Loading data

Loading all the excel files of the data folder:

```{r}
the3files <- path2data |> 
  list.files(full.names = TRUE) |> 
  grep("gadm", x = _, value = TRUE, invert = TRUE) |> 
  set_names2() |> 
  map(read_excel, "Marc", col_types = "text")
```

A function that recode sex:

```{r}
recode_sex <- function(x) {
  sex_val <- sort(unique(x$sex))
  males   <- grep("^M|na", sex_val, TRUE, value = TRUE)
  females <- grep("^M|na", sex_val, TRUE, value = TRUE, invert = TRUE)
  mutate(x, across(sex, ~ recode_values(.x, males ~ "male", females ~ "female")))
}
```

A piece of code common to all the collections:

```{r}
common_data_processing <- function(x) {
  x |> 
    filter(!if_all(everything(), is.na)) |>
    recode_sex() |>
    separate_wider_delim(ID, regex("\\s+"), names = c("sID", "cdate", "plate")) |>
    mutate(across(titer, as.double))
}
```

Combining the last two collections:

```{r}
last2collections <- the3files |> 
  extract(c("Dec2022", "Apr24")) |> 
  map(select, -date_collection) |> 
  bind_rows() |> 
  common_data_processing() |> 
  mutate(collection_date = ymd(paste(year_collection, month_collection, day_collection,
                                     sep = "-"))) |> 
  select(participant, province, hospital, site, collection, sample_ID, age_group, age,
         sex, collection_date, plate, titer, conclusion)
```

Now to the first collection:

```{r}
first_collection <- the3files$DEC2019 |> 
  common_data_processing() |> 
  mutate(date_chr = if_else( grepl("[./]", date_collection), date_collection, NA),
         date_num = if_else(!grepl("[./]", date_collection), date_collection, NA),
         across(date_num, ~ .x |> as.numeric() |> as_date("1899-12-30")),
         across(date_chr, dmy),
         collection_date = coalesce(date_num, date_chr)) |> 
  select(participant, hospital, collection, sample_ID, age_group, age, sex,
         collection_date, plate, titer, conclusion)
```

The total collection:

```{r warning = FALSE}
dataset <- last2collections |>
  bind_rows(first_collection) |> 
  mutate(province = ifelse(is.na(province), hospital, province),
         across(province, ~ recode_values(.x, "AG" ~ "An Giang",
                                              "BD" ~ "Binh Dinh",
                                              "DL" ~ "Dak Lak",
                                              "DT" ~ "Dong Thap",
                                              "HC" ~ "HCMC",
                                              "HU" ~ "Hue",
                                              "KG" ~ "Kien Giang",
                                              "KH" ~ "Khanh Hoa",
                                              "QN" ~ "Quang Ngai",
                                              "ST" ~ "Soc Trang",
                                              "Daklak" ~ "Dak Lak",
                                              "HUE" ~ "Hue",
                                              "HCM" ~ "HCMC", default = .x)),
         year = year(collection_date),
         time_point = recode_values(year, 2019:2020 ~ "T1",
                                          2022:2023 ~ "T2", default = "T3"),
         across(age, ~ ifelse(participant == "0158" & province == "Dak Lak" &
                                collection == "42", "48", .x)),
         age_unit = str_remove_all(age, "\\d"),
         age2 = as.numeric(str_remove_all(age, "\\D")),
         age3 = as.numeric(age),
         age4 = ifelse(is.na(age3), ifelse(age_unit %in% c("M", "th"), age2 / 12,
                                           ifelse(age_unit == "ng", age2 / 365, age2)),
                       age3),
         age = ifelse(age4 > 100, year - age4, age4),
         age_group = cut(age, c(0:20, seq(25, 70, 5), 100), right = FALSE),
         across(conclusion, ~ ifelse(.x == "pós", "Pos", .x))) |> 
  select(-c(year, age_unit, age2, age3, age4, hospital)) |> 
  arrange(time_point, province, age_group, collection_date)
```

A hash table to generate site ID from province names:

```{r}
hash <- dataset |> 
  filter_out(time_point == "T1") |> 
  filter(age > 20) |> 
  select(province, site) |> 
  unique() |> 
  with(setNames(site, province))
```

Completing the site ID:

```{r}
dataset <- mutate(dataset, across(site, ~ ifelse(is.na(.x), hash[province], .x)))
```

Which gives:

```{r}
dataset
```


## Sampling locations

Loading the `sf` object of the provinces of Vietnam:

```{r}
gadm1 <- readRDS(paste0(path2data, "gadm36_VNM_1_sf.rds"))
st_crs(gadm1) <- st_crs(gadm1)
```

Plotting the provinces from which samples where collected:

```{r fig-map, fig.height = 6}
sel <- dataset |>
  pull(province) |> 
  unique() |> 
  str_replace("Hue", "Thua Thien Hue") |> 
  str_replace("HCMC", "Ho Chi Minh")

opar <- par(plt = rep(0:1, 2))

gadm1 |> 
  st_geometry() |> 
  plot(col = "lightyellow")


gadm1 |> 
  filter(VARNAME_1 %in% setdiff(sel, "Kien Giang")) |> 
  st_geometry() |> 
  plot(add = TRUE, col = 2)

gadm1 |> 
  filter(VARNAME_1 == "Kien Giang") |> 
  st_geometry() |> 
  plot(add = TRUE, col = 7)

par(opar)
```


## Overview of the data

Number of sample per site per time point:

```{r}
with(dataset, table(province, time_point))
```

```{r}
saveRDScache(with(dataset, table(province, time_point)), "nb_samples_prov_timepoint.rds")
```

A tuning of the `barplot()` function:

```{r}
barplot2 <- function(x, ...) {
  n1 <- 5; n2 <- 10; ad <- 20
  with(x, {
    barplot(Freq, space = 0, col = 4, ...)
    mtext(province[1], line = -.5)
  })
  segments( 0, n1, ad, n1, col = 2, lwd = 4)
  segments(ad, n2, 31, n2, col = 2, lwd = 4)
}
```

Number of sample per age class, per province, per time point:

```{r fig-nb_samples_site_timepoint_ageclass, fig.width = 7, fig.height = 50 / 3}
tmp <- dataset |> 
  with(table(age_group, province, time_point)) |>
  as.data.frame()

ylim <- c(0, max(tmp$Freq))

opar <- par(mfcol = c(10, 3), plt = c(.105, .97, .1, .9))
tmp |> 
  group_by(time_point, province) |> 
  group_split() |> 
  walk(~ barplot2(.x, ylim = ylim))
par(opar)
```

Number of sample per day and site:

```{r fig-nb_sample_day_site, fig.width = 7, fig.height = 5}
hz_lines <- 1:10

dataset |> 
  group_by(province, collection_date) |> 
  tally() |> 
  ungroup() |> 
  mutate(y = max(hz_lines) + 1 - as.numeric(as.factor(province))) |> 
  with(plot(collection_date, y, cex = sqrt(n), ann = FALSE, axes = FALSE, col = 4,
            xpd = TRUE))

lb <- 2020:2024
axis(1, ymd(paste(lb, "0101")), lb)
abline(h = hz_lines, col = "grey")
text2(rep(ymd(20210701), length(hz_lines)), rev(hz_lines), unique(dataset$province))
```

An alternative representation:

```{r fig-cum_nb_sample, fig.width = 7, fig.height = 5}
plot(NA, xlim = range(dataset$collection_date), ylim = c(0, 600), axes = FALSE,
     xlab = NA, ylab = "cumulative number of samples")
axis(1, ymd(paste(lb, "0101")), lb); axis(2)

dataset |> 
  group_by(province, collection_date) |> 
  tally() |> 
  group_modify(~ mutate(.x, cumn = cumsum(n))) |> 
  group_walk(~ with(.x, lines(collection_date, cumn, type = "s", col = 4)))
```

Yet another representation:

```{r fig-sampling_intensity, fig.width = 7, fig.height = 5}
dataset |> 
  group_by(collection_date) |> 
  tally() |> 
  with(plot(collection_date, n, type = "h", col = "4", xlab = NA,
            ylab = "sampling intensity (/day)"))
```

Gender distribution:

```{r fig-sex_distribution, fig.width = .7 * 7, fig.height = .7 * 50 / 3}
opar <- par(mfcol = c(10, 3), plt = c(.05, .95, .01, .91))
dataset |> 
  with(table(sex, province, time_point)) |> 
  as.data.frame() |> 
  group_by(time_point, province) |> 
  group_split() |> 
  walk(~ with(.x, {
    pie2(Freq, NA, col = c(rgb2(247, 114, 184), rgb2(62, 146, 247)))
    mtext(province[1], line = -.5)
  }))
par(opar)
```

```{r}
dataset |> 
  filter(conclusion == "Neg") |> 
  pull(titer) |> 
  max()

dataset |> 
  filter(conclusion == "Equi") |> 
  pull(titer) |> 
  range()

dataset |> 
  filter(conclusion == "Pos") |> 
  pull(titer) |> 
  min()
```

```{r}
ptiters <- function(df, f = log10, ...) {
  plot(df, type = "n", xlab = "titer (IU)", ylab = "density", axes = FALSE, ...)
  lb1 <- unlist(map(10^(1:5), ~ .x * 1:10))
  lb2 <- 1:5
  axis(1, log10(lb1), NA, tcl =  - .25)
  axis(1, lb2, 10^lb2)
  axis(2)
  df |> 
    group_by(cut(x, c(-Inf, f(c(150, 200)), Inf))) |> 
    group_map(~ pad(.x)) |> 
    walk2(c(2, 7, 3), ~ polygon(.x, col = .y, border = NA))
}
```

```{r}
plot_titers <- function(df, f = log10, ...) {
  df |> 
    pull(titer) |> 
    f() |> 
    density() |> 
    extract(1:2) |> 
    as_tibble() |> 
    ptiters(f, ...)
}
```

Plotting the titers:

```{r fig-distribution_titers}
plot_titers(dataset)
```

```{r}
dataset |> 
  group_by(conclusion) |> 
  tally() |> 
  mutate(prop = 100 * n / sum(n))
```

```{r fig-distribution_titers_timepoints, fig.width = .8 * eps * figw, fig.height = .8 * 3 * eps * figh}
opar <- par(mfrow = c(3, 1), cex = 1, plt = c(.13, .97, .18, .95))
dataset |> 
  group_by(time_point) |> 
  group_walk(~ plot_titers(.x, xlim = c(.5, 4.5), ylim = c(0, .8)))
par(opar)
```

```{r}
dataset |> 
  group_by(time_point, conclusion) |> 
  tally() |> 
  mutate(prop = 100 * n / sum(n)) |> 
  pivot_wider(id_cols = -n, names_from = time_point, values_from = prop)
```


## Titer age

```{r}
estimate_prop <- function(x) {
  slot_names <- c("p", "l", "u")
  x <- with(x, setNames(n, conclusion))
  n <- sum(x)
  if (n < 1) return(setNames(rep(NA_real_, 3), slot_names))
  binom.test(x["Pos"], n) |> 
    extract(c("estimate", "conf.int")) |> 
    unlist() |> 
    setNames(slot_names)
}
```

```{r}
xs <- c(.5:19.5, seq(22.5, 67.5, 5), 75)

p_age_profile <- function(x) {
  plot(xs, rep(.5, length(xs)), ylim = 0:1, type = "n", xaxs = "i", yaxs = "i",
       xlim = c(0, 80),
       xlab = "age classe (years)", ylab = "sero-prevalence")
  abline2(v = seq(0, 70, 5))
  abline2(h = seq(0, 1, .1))
  with(x, {
    arrows(xs, l, xs, u, .02, 90, 3, col = 4)
    points(xs, p, pch = 19, col = 4)
  })
  box(bty = "o")
}
```

```{r}
template <- dataset |> 
  group_by(age_group, conclusion) |> 
  tally() |> 
  ungroup() |> 
  select(-n)

prepare_data <- function(x) {
  x |> 
    group_by(age_group, conclusion) |> 
    tally() |> 
    left_join(x = template, y = _, c("age_group", "conclusion")) |> 
    replace_na(list(n = 0)) |> 
    group_by(age_group) |> 
    group_split() |> 
    map_dfr(estimate_prop)
}
```

```{r}
plot_age_profile <- function(x) {
  x |> 
    prepare_data() |> 
    p_age_profile()
}
```

```{r fig-age_profile, fig.width = 7, fig.height = 5}
plot_age_profile(dataset)
```

```{r fig-age_profile_timepoints, fig.width = .8 * eps * figw, fig.height = .8 * 3 * eps * figh}
opar <- par(mfrow = c(3, 1), cex = 1, plt = c(.13, .97, .18, .95))
dataset |> 
  group_by(time_point) |> 
  group_walk(~ plot_age_profile(.x))
par(opar)
```

```{r}
time_change <- function(x) {
  map(1:nrow(first(x)), function(y) map_dfr(x, ~ .x[y, ]))
}
```

```{r}
a <- dataset |> 
  group_by(time_point) |> 
  group_split() |> 
  map(prepare_data) |> 
  time_change()
```

```{r}
time_trend <- function(x) {
  xs <- 1:3
  col <- 4
  with(x, {
    plot(xs, p, xlim = c(0, 4), ylim = 0:1, type = "n", axes = FALSE, ann = FALSE)
    abline(h = seq(.2, .8, .2), col = "lightgrey")
    arrows(xs, l, xs, u, .05, 90, 3, col = col)
    points(xs, p, col = col, pch = 19)
    box(bty = "o")
  })
}
```

```{r fig-seroprevalence_trends, fig.width = 8.4, fig.height = 1.8}
epsl <- .065
opar <- par(mfrow = c(3, 10), cex = 1, plt = rep(c(epsl, 1 - epsl), 2))
dataset |> 
  filter_out(age_group == last(levels(age_group))) |> 
  group_by(time_point) |> 
  group_split() |> 
  map(prepare_data) |> 
  map(~ .x[-nrow(.x), ]) |> 
  time_change() |> 
  walk(time_trend)
par(opar)
```

```{r}
dataset |> 
  with(plot(age, log10(titer)))
```

```{r}
library(mgcv)
```

```{r}
get_reml <- function(x) {
  as.numeric(x$gcv.ubre)
}
```

```{r}
dev_explained <- function(x) {
  summary(x)$dev.expl
}
```

```{r}
AIC(out[[1]])
```


```{r}
model <- dataset |> 
  mutate(log_titer = log10(titer)) |> 
  gam(log_titer ~ s(age, k = 50), data = _, method = "REML")

data2 <- mutate(dataset, log_titer = log10(titer))

out <- c("tp", "ts", "ds", "cr", "cs", "cc", "bs", "ps", "cp", "re", "gp") |>
  purrr::set_names() |> 
  map(~ gam(log_titer ~ s(age, bs = .x, k = 50), data = data2, method = "REML"))

gam(log_titer ~ s(age, k = 50), data = data2, method = "REML")


k.check(model)

x <- seq(0, 70, le = 512)
y <- model |>
  predict(list(age = x), se.fit = .95) |> 
  as_tibble()
plot(x, y$fit, type = "l", ylim = range(y))
lines(x, y$ll)
lines(x, y$ul)
abline(h = log10(c(150, 200)))
```


