---
title: "Measles serology"
number-sections: true
format:
  html:
    toc: true
    toc-depth: 4
editor: source
editor_options: 
  chunk_output_type: console
#bibliography: references.bib
#csl: the-american-naturalist.csl
---

```{r include = FALSE, eval = FALSE}
# The code that adds the .nojekyll file to the GitHub repository:
file <- ".nojekyll"
file.create(file)
gert::git_add(file)
gert::git_commit("Adding the .nojekyll file")
gert::git_push()
rm(file)
```

```{r include = FALSE, eval = FALSE}
run_all <- function() {
  knitr::purl("index.qmd", "tmp.R", documentation = FALSE)
  source("tmp.R")
  file.remove("tmp.R")  
}

rerun_all <- function() {
  rm(list  = grep("run_all", ls(envir = .GlobalEnv), value = TRUE, invert = TRUE),
     envir = .GlobalEnv)
  run_all()
}
```

```{r include = FALSE, purl = FALSE}
setHook("plot.new", NULL, action = "replace")

par3 <- function(...) par(..., mgp = c(1.5, .5, 0), bty = "n")

knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par3(plt = c(.105, .97, .15, .95)) else NULL
  })

figw <- 7  # default is 7 and maximum possible is 8.3
figh <- 5  # default is 5
eps <- .8

knitr::opts_chunk$set(margin     = TRUE,
                      fig.retina = 2,
                      fig.align  = "center",
                      fig.width  = eps * figw,
                      fig.height = eps * figh) 
```


## Constants

The path to the data folder:

```{r}
if (grepl("arc|hoisy", Sys.getenv("USER"))) {
  path2data <- paste0("~/Library/CloudStorage/OneDrive-OxfordUniversityClinical",
                      "ResearchUnit/GitHub/choisy/measles serology/")
}
```

```{r include = FALSE}
path2cache <- paste0(path2data, "cache/")
if (! dir.exists(path2cache)) dir.create(path2cache)
make_path <- function(x) paste0(path2cache, x)
file_exists <- function(x) file.exists(make_path(x))
readRDScache <- function(x) readRDS(make_path(x))
saveRDScache <- function(object, file) saveRDS(object, make_path(file))
```


## Packages

Required packages:

```{r}
required <- c("sf", "readxl", "stringr", "stringi", "purrr", "tidyr", "magrittr",
              "RColorBrewer", "lubridate", "mgcv", "dplyr")
```

Installing those that are not installed yet:

```{r}
to_install <- required[! required %in% installed.packages()[, "Package"]]
if (length(to_install)) install.packages(to_install)
```

Loading the packages for interactive use:

```{r warning = FALSE, message = FALSE}
invisible(sapply(required, library, character.only = TRUE))
```


## Functions

A function that generates names of a vector of character strings from its content:

```{r}
set_names2 <- function(x) {
  setNames(x, x |>
                str_remove("^.*_") |> 
                str_remove("^.*-") |> 
                str_remove(".xlsx$"))
}
```

A tuning of the `text()` function that includes white background to the text:

```{r}
text2 <- function(x, y, labels, cex = 1, bg = "white", padx = 10, pady = .1, ...) {
  w <- strwidth(labels, cex = cex)
  h <- strheight(labels, cex = cex)
  rect(x - w / 2 - padx, y - h / 2 - pady / 2, x + w / 2 + padx, y + h / 2 + pady / 2,
       col = bg, border = NA)
  text(x, y, labels, cex = cex, ...)
}
```

Tuning of the `rgb()` function:

```{r}
rgb2 <- function(...) rgb(..., maxColorValue = 255)
```

Tunings of the `plot()` function:

```{r}
plot2 <- function(...) plot(..., axes = FALSE)
plot3 <- function(...) plot2(..., ann = FALSE)
plot4 <- function(...) plot3(..., type = "n")
```

A tuning of the `pie()` function:

```{r}
pie2 <- function(x, ...) if (all(x == 0)) plot4(1) else pie(x, ...)
```

A function that pads a data frame of coordinates:

```{r}
pad <- function(x) {
  x <- arrange(x, 1)
  xs <- x[[1]]
  tibble(c(first(xs), xs, last(xs)), c(0, x[[2]], 0)) |> 
    setNames(names(x)[1:2])
}
```

Some tunings of the `abline()` function:

```{r}
abline2 <- function(...) abline(..., col = "lightgrey")
abline3 <- function(...) abline(..., col = "grey")
```

Some tunings of the `mgcv::gam()` function:

```{r}
gam2 <- function(...) mgcv::gam(..., method = "REML")
gam3 <- function(formula, data, ...) gam2(formula, data, family = binomial, ...)
```

A function that retrieves the REML from a GAM model:

```{r}
get_reml <- function(x) as.numeric(x$gcv.ubre)
```

A function that retrieves the percentage of deviance explained from a GAM model:

```{r}
dev_explained <- function(x) summary(x)$dev.expl
```

A tuning of the `segments()` function:

```{r}
segments2 <- function(...) segments(..., col = 2, lwd = 4)
```

A function that extends the `tally()` function by adding a column of the corresponding
proportions:

```{r}
tally2 <- function(x) {
  x |> 
    tally() |> 
    mutate(prop = 100 * n / sum(n))
}
```

A tuning of the `arrows()` function:

```{r}
arrows2 <- function(...) arrows(..., angle = 90, code = 3)
```

A tuning of the `points()` function:

```{r}
points2 <- function(...) points(..., pch = 19)
```

A tuning of the `par()` function:

```{r}
par2 <- function(...) par(..., cex = 1)
```

A function that adds the $y$-axis on a log scale:

```{r}
logaxis <- function(side, po10 = 1:5) {
  axis(side, log10(unlist(map(10^(po10), ~ .x * 1:10))), NA, tcl = - .25)
  axis(side, po10, 10^po10)
}
```

A tuning of the `binom.test()` function:

```{r}
binom_test2 <- function(x, n, ...) {
  if (n < 1) return(list(estimate = NA, conf.int = rep(NA, 2)))
  binom.test(x, n, ...)
}
```

A tuning of the `predict()` function:

```{r}
predict2 <- function(...) mgcv::predict.gam(..., type = "response")
```

The following function takes a vector `x` of densities of probabilities as an input and
returns a threshold value above which the values of `x` belong to the `p` confidence
interval:

```{r}
ci2th <- function(v, p) {
  denominator <- sum(v)
  f <- function(th) sum(v[v > th]) / denominator
  optimize(function(y) abs(f(y) - p), range(v))$min
}
```

The following function converts a vector `x` of densities of probabilities into a
vector of quantile belonging where the quantiles are defined by the vector `breaks`:

```{r}
quantval2quantcat <- function(x, breaks = seq(.95, .05, -.05)) {
  as.numeric(cut(x, c(0, map_dbl(breaks, ci2th, v = x), max(x))))
}
```

A function that extracts the last `x` characters of a character string `x`:

```{r}
#lastx <- function(v, x) {
#  nb <- nchar(v)
#  substr(v, nb - x + 1, nb)
#}
```

A function that convert Vietnamese characters into Latin characters:

```{r}
convert <- function(x) stringi::stri_trans_general(x, "Latin-ASCII")
```

A function that convert the names of a data frame into Latin ASCII:

```{r}
latin_var_names <- function(df) {
  names(df) <- convert(names(df))
  df
}
```

A tuning of the `lubridate::as_date` function:

```{r}
as_date_excel <- function(...) lubridate::as_date(..., origin = "1899-12-30")
```

A function that converts a date from number of days in character string into an R date
object, using Excel's origin of time:

```{r}
convert_date <- function(x) x |> as.numeric() |> as_date_excel()
```

A tuning of the `readxl::read_excel()` function:

```{r}
read_excel2 <- function(...) readxl::read_excel(..., col_types = "text")
```

Tunings of the `min()` and `max()` functions:

```{r}
min2 <- function(...) min(..., na.rm = TRUE)
max2 <- function(...) max(..., na.rm = TRUE)
```

A function that replaces all the `NA` of a data frame by 0:

```{r}
replace_na2 <- function(x) mutate(x, across(everything(), ~ replace_na(.x, 0)))
```

A function that plots an `sf` object using one of its columns to define the colors:

```{r}
plot_sf2 <- function(sf, col) {
  sf |> 
    st_geometry() |> 
    plot(col = sf[[col]])
}
```

A function that adds the column-wise sum of the numeric variables at the bottom of a
data frame:

```{r}
add_colsum <- function(df) {
  bind_rows(df, map_dfr(df, function(x) ifelse(is.numeric(x), sum(x), NA)))
}
```


## Data

### Functions

A hash table to recode sex:

```{r}
hash_sex <- setNames(rep(c("female", "male"), c(4, 3)),
                     c("f", "nu", "nnu", "nuu", "m", "n", "nam"))
```

A piece of code common to all the sample collections:

```{r}
common_data_processing <- function(x) {
  x |> 
    filter_out(if_all(everything(), is.na)) |>
    mutate(across(sex, ~ hash_sex[tolower(convert(.x))])) |> 
    separate_wider_delim(ID, regex("\\s+"), names = c("sID", "cdate", "plate")) |>
    mutate(across(titer, as.double))
}
```

A function that generates age groups from age:

```{r}
make_age_group <- function(x) cut(x, c(0:20, seq(25, 70, 5), 100), right = FALSE)
```


### Data

#### South

Loading all the excel files of the data folder:

```{r}
the3files <- paste0(path2data, "south") |> 
  list.files(full.names = TRUE) |> 
  set_names2() |> 
  map(read_excel2, "Marc")
```

The first collection:

```{r}
first_collection <- the3files$DEC2019 |> 
  common_data_processing() |> 
  mutate(date_chr = if_else( grepl("[./]", date_collection), date_collection, NA),
         date_num = if_else(!grepl("[./]", date_collection), date_collection, NA),
         across(date_num, ~ .x |> as.numeric() |> as_date_excel()),
         across(date_chr, dmy),
         collection_date = coalesce(date_num, date_chr)) |> 
  select(participant, hospital, collection, sample_ID, age_group, age, sex,
         collection_date, plate, titer, conclusion)
```

Combining the collections:

```{r warning = FALSE}
south <- the3files |> 
  extract(c("Dec2022", "Apr24")) |> 
  map(select, -date_collection) |> 
  bind_rows() |> 
  common_data_processing() |> 
  mutate(collection_date = ymd(paste(year_collection, month_collection, day_collection,
                                     sep = "-"))) |> 
  select(participant, province, hospital, site, collection, sample_ID, age_group, age,
         sex, collection_date, plate, titer, conclusion) |> 
  bind_rows(first_collection) |> 
  mutate(province = ifelse(is.na(province), hospital, province),
         across(province, ~ recode_values(.x, "AG"     ~ "An Giang",
                                              "BD"     ~ "Binh Dinh",
                                              "DL"     ~ "Dak Lak",
                                              "DT"     ~ "Dong Thap",
                                              "HC"     ~ "HCMC",
                                              "HU"     ~ "Hue",
                                              "KG"     ~ "Kien Giang",
                                              "KH"     ~ "Khanh Hoa",
                                              "QN"     ~ "Quang Ngai",
                                              "ST"     ~ "Soc Trang",
                                              "Daklak" ~ "Dak Lak",
                                              "HUE"    ~ "Hue",
                                              "HCM"    ~ "HCMC", default = .x)),
         year = year(collection_date),
         time_point = recode_values(year, 2019:2020 ~ "T1",
                                          2022:2023 ~ "T2", default = "T3"),
         across(age, ~ ifelse(participant == "0158" & province == "Dak Lak" &
                                collection == "42", "48", .x)),
         age_unit = str_remove_all(age, "\\d"),
         age2 = as.numeric(str_remove_all(age, "\\D")),
         age3 = as.numeric(age),  # note: this generate NAs
         age4 = ifelse(is.na(age3), ifelse(age_unit %in% c("M", "th"), age2 / 12,
                                           ifelse(age_unit == "ng", age2 / 365, age2)),
                       age3),
         age = ifelse(age4 > 100, year - age4, age4),
         age_group = make_age_group(age),
         across(conclusion, ~ ifelse(.x == "pós", "Pos", .x))) |> 
  select(-c(year, age_unit, age2, age3, age4, hospital)) |> 
  arrange(time_point, province, age_group, collection_date)
```

A hash table to generate site ID from province names:

```{r}
hash_id <- south |> 
  filter_out(time_point == "T1") |> 
  filter(age > 20) |> 
  select(province, site) |> 
  unique() |> 
  with(setNames(site, province))
```

Completing the site ID when missing:

```{r}
south <- mutate(south, across(site, ~ ifelse(is.na(.x), hash_id[province], .x)))
```

Which gives:

```{r}
south
```


#### North

Hash table for conclusion:

```{r}
hash_conclusion <- c("Am tinh"    = "Neg",
                     "Duong tinh" = "Pos",
                     "Nghi ngo"   = "Equi",
                     "Equi"       = "Equi",
                     "Pos"        = "Pos",
                     "Neg"        = "Neg",
                     "pós"        = "Pos")
```

Hash table for the variable names of the northern part of the country:

```{r}
hash_translation <- c("Benh vien"     = "hospital",  
                      "Ten benh vien" = "province",
                      "Lab ID"        = "Lab ID",
                      "Ma benh vien"  = "site",
                      "Dot lay mau"   = "collection",
                      "Ma benh pham"  = "participant",
                      "Study ID"      = "sample_ID",
                      "Ten viet tat"  = "initials",
                      "Do tuoi"       = "age",
                      "Gioi tinh"     = "sex",
                      "Khoa"          = "ward",
                      "Ngay nhan mau" = "reception_date",
                      "Ngay thu mau"  = "collection_date",
                      "Note"          = "Note",
                      "Ghi chu"       = "Note2")
```

A function that translates the names of the variable of a data frame from Vietnamese to
English using the `hash_translation` hash table:

```{r}
convert_var_names <- function(df) {
  names(df) <- hash_translation[names(df)]
  df
}
```

Putting the titers data into shape:

```{r}
titers <- map(c("IgG Serion 2025 Measles 3021 mau.xls",
                "IgG Serion 2025 Measles 663 mau.xls"),
              ~ read_excel(paste0(path2data, "north/titers/", .x))) |> 
  bind_rows() |> 
  latin_var_names() |> 
  select(-STT, -`ID mau`) |> 
  rename(titer      = `Nong do khang the`,
         conclusion = `Ket qua`) |> 
  mutate(across(conclusion, ~ hash_conclusion[convert(.x)])) |> 
  unique()
```

Generating the data set for the northern part of the country:

```{r warning = FALSE}
north <- c("02FL Tracking sample round 1 to 6 Final.xlsx",
           "02FL Tracking sample round 7 to 9.xlsx") |> 
  rep(2:1) |> 
  map2(c(1, 2, 1), ~ read_excel2(paste0(path2data, "north/metadata/", .x), .y)) |> 
  bind_rows() |> 
  latin_var_names() |> 
  convert_var_names() |> 
  mutate(across(province, convert),
         across(province, ~ .x |>
                  str_replace("Viet Tiep" , "Hai Phong") |> 
                  str_replace("Thanh Nhan", "Ha Noi")),
         across(collection_date, convert_date),
         across(reception_date, ~ as_date(ifelse(.x == "30/10/2020", ymd(20201030),
                                                 convert_date(.x)))),
         sex       = hash_sex[tolower(sex)],
         unit      = convert(str_remove_all(age, "\\d")),
         age       = as.numeric(str_remove_all(age, "\\D")),
         age       = ifelse(unit == "ngay", age / 365,
                            ifelse(unit == "thang", age / 12, age)),
         age_group = make_age_group(age)) |> 
  left_join(titers, y = _, "Lab ID") |> 
  rename(lab_ID = `Lab ID`) |> 
  select(participant, province, site, collection, sample_ID, age_group, age, sex,
         collection_date, reception_date, lab_ID, titer, conclusion) |> 
  mutate(time_point = ifelse(collection == "01", "T1", "T2"))
```

Check:

```{r}
map_int(north, ~ .x |> is.na() |> sum())
```

```{r}
north |> 
  group_by(collection) |> 
  summarize(from_recep      = min2(reception_date),
            to_recep        = max2(reception_date),
            n               = n(),
            NA_coll         = sum(is.na(collection_date)))
```

Checking inconsistency between dates of collection and reception:

```{r}
filter(north, reception_date < collection_date)
```

#### All

```{r}
dataset <- north |> 
  mutate(collection_date = as_date(ifelse(is.na(collection_date),
                                          reception_date, collection_date))) |> 
  select(-reception_date, -lab_ID) |> 
  bind_rows(select(south, -plate)) |> 
  filter_out(is.na(age_group))
```

Checking the thresholds:

```{r}
dataset |> 
  filter(conclusion == "Neg") |> 
  pull(titer) |> 
  max()

dataset |> 
  filter(conclusion == "Equi") |> 
  pull(titer) |> 
  range()

dataset |> 
  filter(conclusion == "Pos") |> 
  pull(titer) |> 
  min()
```

Number of samples per site per time point:

```{r}
(nb_samples <- dataset |> 
  group_by(province, time_point) |> 
  tally() |> 
  ungroup() |> 
  pivot_wider(names_from = time_point, values_from = n) |> 
  replace_na2())
```

```{r echo = FALSE}
saveRDScache(nb_samples, "nb_samples_prov_timepoint.rds")
```

## Sampling locations

Loading the and modifying `sf` object of the provinces of Vietnam:

```{r message = FALSE}
col_map <- c(brewer.pal(3, "Set1"), "lightyellow")

gadm1 <- readRDS(paste0(path2data, "gadm36_VNM_1_sf.rds"))
st_crs(gadm1) <- st_crs(gadm1)

gadm1 <- gadm1 |> 
  rename(province = VARNAME_1) |> 
  select(province, geometry) |> 
  mutate(across(province, ~ .x |>
                  str_replace("Thua Thien Hue", "Hue") |>
                  str_replace("Ho Chi Minh", "HCMC")),
         centroid = geometry |> st_centroid() |> st_coordinates()) |> 
  left_join(nb_samples, "province") |> 
  replace_na2() |> 
  mutate(col = col_map[(T1 < 1) + (T2 < 1) + (T3 < 1) + 1])
```

Plotting the provinces from which samples where collected:

```{r fig-map, fig.width = .7 * 7, fig.height = 6}
#| fig-cap: "The provinces of Vietnam from which serum samples are collected for the
#| first time point only (Dec 2019 in green), the first two time points only (Dec 2019
#| and Dec 2022 in blue), and for the three time points (Dec 2019, Dec 2022 and Apr
#| 2024 in red)."
opar <- par(plt = rep(0:1, 2))
plot_sf2(gadm1, "col")
par(opar)
```


## Samples metadata

### Sampling intensity

Number of samples per day and site:

```{r fig-nb_sample_day_site, fig.height = 1.2 * eps * figh}
#| fig-cap: "The number of samples collected as a function of time and space
#| (proportional to the areas of the circles). The provinces are arranged from bottom
#| to top according to the latitude of the centroid of their administrative polygon and
#| the color code is the same as for @fig-map."
hzl <- 1:20

opar <- par(plt = c(.105, .97, .1, .9))

nb_samples2 <- gadm1 |>
  st_drop_geometry() |> 
  mutate(latitude = centroid[, "Y"]) |> 
  select(province, latitude, col) |> 
  left_join(nb_samples, y = _, "province") |> 
  arrange(desc(latitude))

dataset |> 
  group_by(province, collection_date) |> 
  tally() |> 
  ungroup() |> 
  mutate(y = max(hzl) + 1 - as.numeric(factor(province, nb_samples2$province))) |> 
  left_join(nb_samples2, "province") |> 
  slice_sample(prop = 1) |> 
  with(plot3(collection_date, y, cex = sqrt(n), xpd = TRUE, col = col))

lb <- 2020:2024
axis(1, ymd(paste(lb, "0101")), lb)
abline3(h = hzl)
text2(rep(ymd(20210701), length(hzl)), rev(hzl), nb_samples2$province)

par(opar)
```

Number of samples:

```{r}
nb_samples2 |> 
  select(-c(latitude, col)) |> 
  mutate(total = T1 + T2 + T3) |> 
  add_colsum() |> 
  print(n = Inf)
```

An alternative representation:

```{r fig-sampling_intensity, fig.width = eps * figw * .865 / .83}
#| fig-cap: "The sampling intensity (grey, left y-axis) and the cumulative number of
#| samples as a function of space and time. The the latter, the color code is the same
#| as for @fig-map."
opar <- par(plt = c(.0825, .9125, .15, .95))
dataset |> 
  group_by(collection_date) |> 
  tally() |> 
  with(plot(collection_date, n, type = "h", col = "grey", xlab = NA,
            ylab = "sampling intensity (nb samples/day)"))

par(new = TRUE)

plot2(NA, xlim = range(dataset$collection_date), ylim = c(0, 600), ann = FALSE)
axis(4)
mtext("cumulative number of samples", 4, 1.5)

dataset |> 
  left_join(select(nb_samples2, province, col), "province") |> 
  group_by(province, collection_date) |> 
  summarise(n = n(), col = unique(col), .groups = "drop_last") |> 
  group_modify(~ mutate(.x, cumn = cumsum(n))) |> 
  group_walk(~ with(.x, lines(collection_date, cumn, type = "s", col = col)))

par(opar)
```


### Age distributions

A tuning of the `barplot()` function:

```{r}
barplot2 <- function(x, n1 = 5, n2 = 10, ad = 20, mn = 0, mx = 31, col = 4, ...) {
  with(x, {
    barplot(Freq, space = 0, col = col, ...)
    mtext(province[1], line = -.5)
  })
  segments2(mn, n1, ad, n1)
  segments2(ad, n2, mx, n2)
}
```

Number of sample per age class, per province, per time point:

```{r fig-nb_samples_site_timepoint_ageclass, fig.width = 7, fig.height = 50 / 3}
tmp <- south |> 
  with(table(age_group, province, time_point)) |>
  as.data.frame()

ylim <- c(0, max(tmp$Freq))

opar <- par2(mfcol = c(10, 3), plt = c(.12, .985, .1, .9))
tmp |> 
  group_by(time_point, province) |> 
  group_split() |> 
  walk(~ barplot2(.x, ylim = ylim))
par(opar)
```


### Gender distribution

```{r fig-sex_distribution, fig.width = .7 * 7, fig.height = .7 * 50 / 3}
opar <- par(mfcol = c(10, 3), plt = c(.05, .95, .01, .91))
south |> 
  with(table(sex, province, time_point)) |> 
  as.data.frame() |> 
  group_by(time_point, province) |> 
  group_split() |> 
  walk(~ with(.x, {
    pie2(Freq, NA, col = c(rgb2(247, 114, 184), rgb2(62, 146, 247)))
    mtext(province[1], line = -.5)
  }))
par(opar)
```


## Titers

A function that plots a 3-color distribution from a 2-column data frame `df` that
contains the coordinates of the distribution, with $x$-values in the first column
named `x` and $y$-values in the second column, on a decimal log scale of the $x$
values:

```{r}
ptiters <- function(df, main = NULL, th1 = 150, th2 = 200, lb2 = 1:5, ...) {
  plot2(df, type = "n", xlab = "titer (IU/L)", ylab = "density", ...)
  logaxis(1); axis(2)
  df |> 
    group_by(cut(x, c(-Inf, log10(c(th1, th2)), Inf))) |> 
    group_map(~ pad(.x)) |> 
    walk2(c(2, 7, 3), ~ polygon(.x, col = .y, border = NA))
  if (! is.null(main)) title(main, line = -.25, cex.main = 1)
}
```

A function that (1) generates the 2-column input data frame of the function
`ptiters()` from the column named `titer` of the data frame `df` and then (2) feeds it
to the `ptiters()` function to draw the 3-color distribution:

```{r}
plot_titers <- function(df, main = NULL, ...) {
  df |> 
    pull(titer) |> 
    log10() |> 
    density() |> 
    extract(1:2) |> 
    as_tibble() |> 
    ptiters(main, ...)
}
```

Plotting the distribution of all the titers, across the 3 time points and the 10 sites:

```{r fig-distribution_titers}
plot_titers(south)
```

The corresponding proportions in each of the 3 categories of titer values are the
following:

```{r}
south |> 
  group_by(conclusion) |> 
  tally2()
```

The distributions of all the titers across the 10 sites for each time point separately:

```{r fig-distribution_titers_timepoints, fig.width = .7 * eps * figw, fig.height = .7 * 3 * eps * figh}
timepoints <- c(paste("Dec", c(2019, 2022)), "Apr 2024")

opar <- par2(mfrow = c(3, 1), plt = c(.13, .97, .18, .95))
south |> 
  group_by(time_point) |> 
  group_split() |> 
  walk2(timepoints, ~ plot_titers(.x, .y, xlim = c(.5, 4.5), ylim = c(0, .8)))
par(opar)
```

The corresponding proportions in each of the 3 categories of titer values and for each
time point are the following:

```{r}
south |> 
  group_by(time_point, conclusion) |> 
  tally2() |> 
  pivot_wider(id_cols = -n, names_from = time_point, values_from = prop)
```


### Age profile

A function that adds the thresholds values to a plot:

```{r}
add_thresholds <- function(ths = c(150, 200), col = 2:3) {
  abline(h = log10(ths), col = col, lty = 2, lwd = 2)
}
```

A function that plots an image of the distribution of titers (along the $y$ axis) for
all the age classes (on the $x$ axis):

```{r}
image_titer_density_age <- function(data, xaxis = TRUE, xlab = "age (years)", n = 512,
                                    min_titer = NULL, max_titer = NULL, buff = TRUE,
                                    breaks = seq(.95, .05, -.05)) {
  if (is.null(min_titer)) min_titer <- min(data$titer)
  if (is.null(max_titer)) max_titer <- max(data$titer)

  if (buff) {
    bf <- .04 * (max_titer - min_titer)
    min_titer <- min_titer - bf
    max_titer <- max_titer + bf
  }
  
  data |> 
    filter(age < 70) |> 
    group_by(age_group) |> 
    group_map(~ density(.x$titer, from = min_titer, to = max_titer, n = n)$y) |> 
    rep(c(rep(1, 20), rep(5, 10))) |> # age classes
    map(quantval2quantcat, breaks) |> # this computes the categorical HDI
    as.data.frame() |> 
    as.matrix() |> 
    t() |>
    image(1:70 - .5, seq(min_titer, max_titer, le = n), z = _, axes = FALSE,
          xlab = xlab, ylab = "titer (IU/L)",
          col = hcl.colors(length(breaks), "YlOrRd", rev = TRUE))
  
  if (xaxis) axis(1)
  logaxis(2)
  add_thresholds()
  box(bty = "o")
}
```

A function that adds an axis of the ages on the top of a figure:

```{r}
add_years_on_top <- function(at = 10 * 0:7, line = 1.7) {
  axis(3, at)
  mtext("age (years)", line = line)
}
```

Let's plot the distributions of the titers as a function of age:

```{r fig-titer_density_age, fig.height = 2 * eps * figh}
opar <- par(plt = c(.105, .97, .5, 1 - (1 - .95) / 2 - .15 / 2))
south |> 
  filter_out(age_group == "[70,100)") |> 
  with(plot(age, log10(titer), col = adjustcolor(4, .2), xlim = c(0, 70), axes = FALSE,
            xlab = NA, ylab = "titer (IU/L)", xaxs = "i", xpd = TRUE))
logaxis(2); box(bty = "o")
add_years_on_top()
add_thresholds()
par(plt = c(.105, .97, .15 / 2 + (1 - .95) / 2, .5), new = TRUE)
south |>
  mutate(across(titer, log10)) |> 
  image_titer_density_age()
par(opar)
```


## Sero-prevalence

This function takes a data frame that contains at least 2 columns named `age_group` and
`titer`) and estimates sero-prevalence with confidence interval per `age_group` where
positivity is defined by a `titer` higher than `th`:

```{r}
prepare_data <- function(df, th) {
  df |> 
    with(table(age_group, titer > th)) |> 
    as.data.frame() |> 
    pivot_wider(names_from = Var2, values_from = Freq) |> 
    mutate(tmp = map2(`TRUE`, `TRUE` + `FALSE`, binom_test2) |> 
             map(extract, c("estimate", "conf.int")) |>
             map(unlist) |> 
             map(setNames, c("p", "l", "u"))) |> 
    unnest_wider(tmp)
}
```

A tuning of the `plot()` function:

```{r}
plotn <- function(...) {
  plot(..., type = "n", xaxs = "i", yaxs = "i", ylab = "sero-prevalence")
}
```

This function plots the output of `prepare_data()`:

```{r}
make_p_age_profile <- function(xs = c(.5:19.5, seq(22.5, 67.5, 5), 75)) {
  function(x, main = NULL) {
    plotn(xs, rep(.5, length(xs)), ylim = 0:1, xlim = c(0, 80),
          xlab = "age classe (years)")
    abline2(v = seq(0, 70, 5))
    abline2(h = seq(0, 1, .1))
    with(x, {
      arrows2(xs, l, xs, u, .02, col = 4)
      points2(xs, p, col = 4, xpd = TRUE)
    })
    box(bty = "o")
    if (! is.null(main)) title(main, line = -9, cex.main = 1)
  }
}

p_age_profile <- make_p_age_profile()
```

A wrapper around the two previous functions:

```{r}
plot_age_profile <- function(df, th, main = NULL) {
  df |> 
    prepare_data(th) |> 
    p_age_profile(main)
}
```

Sero-prevalence profile across the 3 time points and the 10 sites:

```{r}
plot_age_profile(south, 200)
```

Sero-prevalence profiles across the 10 sites for each of the time points:

```{r fig.width = .7 * eps * figw, fig.height = .7 * 3 * eps * figh}
opar <- par2(mfrow = c(3, 1), plt = c(.13, .97, .18, .95))
south |> 
  group_by(time_point) |> 
  group_split() |> 
  walk2(timepoints, ~ plot_age_profile(.x, 200, .y))
par(opar)
```

The following function returns a list of data frames of 3 columns too where the number
of slots of the list and the number of rows of the data frames are switched.

```{r}
time_change <- function(x) {
  map(1:nrow(first(x)), function(y) map_dfr(x, ~ .x[y, ]))
}
```

This function takes a 3-column (proportion estimate and confidence interval bounds) /
3-row (for 3 time points) data frame that corresponds to a given age group and plots 
the change in sero-prevalence as a function of time.

```{r}
time_trend <- function(x, xs = 1:3, col = 4) {
  with(x, {
    plot4(xs, p, xlim = c(0, 4), ylim = 0:1)
    abline2(h = seq(.2, .8, .2))
    arrows2(xs, l, xs, u, .05, col = col)
    points2(xs, p, col = col)
    box(bty = "o")
  })
}
```

Showing the change in seroprevalence as a function of time for all the age groups
separately:

```{r fig.width = 8.4, fig.height = 1.8}
epsl <- .065
opar <- par(mfrow = c(3, 10), plt = rep(c(epsl, 1 - epsl), 2))
south |> 
  filter_out(age_group == last(levels(age_group))) |> 
  group_by(time_point) |> 
  group_split() |> 
  map(prepare_data, 200) |> 
  map(~ .x[-nrow(.x), ]) |> 
  time_change() |> 
  walk(time_trend)
par(opar)
```


### GAM

#### Functions

A tuning of the `predict2()` function:

```{r}
predict3 <- function(model, x = seq(0, 70, le = 512), ci = .95) {
  bind_cols(tibble(x  = x),
            model |> 
              predict2(list(age = x), se.fit = ci) |> 
              as_tibble())
}
```

A function that adds prediction and confidence interval to a plot:

```{r}
add_ribbon <- function(pred, col = 4) {
  with(pred, {
    polygon(c(x, rev(x)), c(ll, rev(ul)), border = NA, col = adjustcolor(col, .2))
    lines(x, fit, col = col, lwd = 2)
  })
}
```

where `pred` is an output of the `predict3()` function and a 4-column data frame with
names `x`, `fit`, `ll` and `ul`. This function plots the predictions of a GAM model:

```{r}
plot_pred <- function(pred, ylim = NULL, col = 4, ...) {
  if (is.null(ylim)) ylim <- 0:1
  with(pred, plotn(x, fit, ylim = ylim, xlab = "age (years)", ...))
  abline2(v = c(.75, 1.5, seq(5, 65, 10)))
  abline3(v = seq(10, 60, 10))
  add_ribbon(pred, col)
  box(bty = "o")
}
```

where `pred` is an output of the `predict3()` function. A wrapper around the
`predict3()`, `plot_pred()` and `add_ribbon()` functions:

```{r}
plot_seroprevalences <- function(data, bs = "tp", k = 50, th1 = 200, th2 = 150) {
  gam3((titer > th1) ~ s(age, bs = bs, k = k), data) |> 
    predict3() |> 
    plot_pred(col = 3)
  
  gam3((titer > th2) ~ s(age, bs = bs, k = k), data) |> 
    predict3() |> 
    add_ribbon(2)
  
  abline3(h = 1:9 / 10)
}
```


#### Analysis

A function that runs 10 different smoothing bases on a GAM model and compare them:

```{r}
compare_bs <- function(data, th = 200, k = 50, n = 1000) {
  tibble(
    smoothing_basis = c("thin plate", "thin plate", "Duchon", "cubic regression",
                        "cubic regression", "cyclic cubic", "B-spline", "P-spline",
                        "cyclic P-spline", "Gaussian process"),
    code = c("tp", "ts", "ds", "cr", "cs", "cc", "bs", "ps", "cp", "gp")) |> 
    mutate(model   = map(code, ~ gam3((titer > th) ~ s(age, bs = .x, k = k), data)),
           REML    = map_dbl(model, get_reml),
           per_dev = map_dbl(model, dev_explained),
           AIC     = map_dbl(model, AIC),
           k_check = map(model, ~ as_tibble(k.check(.x, n.rep = n)))) |> 
    unnest_wider(k_check)
}
```

Let's try 10 smoothing bases to see whether one stands out as much better that the
others:

```{r eval = FALSE}
models <- compare_bs(south)
```

```{r include = FALSE}
if (file_exists("models.rds")) {
  models <- readRDScache("models.rds")
} else {
  models <- compare_bs(south)
  saveRDScache(models, "models.rds")
}
```

which gives:

```{r}
select(models, -model)
```

Let's then stick to the thin plate smoothing. Plotting smoothed trends:

```{r fig-gam_seroprev}
models |> 
  filter(code == "tp") |> 
  pull(model) |> 
  first() |>
  predict3() |>
  plot_pred()
abline3(h = 1:9 / 10)
```

With the two thresholds together:

```{r fig-gam_seroprev2}
plot_seroprevalences(south)
```

And together with the distributions of titers as a function of age:

```{r fig-titers_seroprevalences, fig.height = 2 * eps * figh}
opar <- par(plt = c(.105, .97, .5, 1 - (1 - .95) / 2 - .15 / 2))
south |> 
  mutate(across(titer, log10)) |> 
  image_titer_density_age(xaxis = FALSE, xlab = NA)
add_years_on_top()
par(plt = c(.105, .97, .15 / 2 + (1 - .95) / 2, .5), new = TRUE)
plot_seroprevalences(south)
par(opar)
```


## TO DO



* run the gam2 model only on the range of age where we have (enough) data

```{r}
subset_vector <- function(x, th) x[x > th]
```

```{r}
smoother <- function(df, bs = "tp", k = -1) {
  c("p", "l", "u") |> 
    setNames(c("fit", "ll", "ul")) |> 
    map(~ gam2(reformulate(paste0('s(age, bs="', bs, '", k=', k, ")"), .x), betar, df))
}
```

```{r}
gam_predictions <- function(data, th = 200, n = 512, bs = "tp", k = -1) {
  data <- data |> 
    mutate(tmp = age_group |> str_remove("\\[") |> str_remove("\\)")) |> 
    separate(tmp, c("l", "u")) |> 
    mutate(across(c(l, u), as.integer))
  
  ndata <- data |> 
    select(l, u) |> 
    range() |> 
    c(n) |> 
    setNames(c("from", "to", "length")) |> 
    as.list() |> 
    do.call(seq, args = _)
  
  data |> 
    group_by(age_group) |> 
    summarise(age = mean(age)) |> 
    left_join(na.exclude(prepare_data(data, th)), "age_group") |> 
    smoother(bs, k) |> 
    map(predict2, newdata = list(age = ndata)) |> 
    as_tibble() |> 
    bind_cols(tibble(x = ndata)) |> 
    relocate(x, .before = 1)
}
```

```{r}
gam_predictions(south)
```

```{r}
south |> 
  filter(province == "An Giang", time_point == "T1") |> 
  filter_out(age < 14 | age > 70) |>
  gam_predictions() |> 
  plot_pred()
```

###

* Model probability of seropositivity between the two thresholds

```{r}
titer2protection <- function(x) {
  ifelse(x < 150, 0, ifelse(x > 200, 1, (x - 150) / 50))
}
```

```{r}
south |> 
  mutate(titer2 = titer2protection(titer)) |> 
  pull(titer2) |> 
  mean()
```

```{r}
?dbeta
```

```{r}
library(MASS)
```


```{r}
th <- 150
#prepare_data <- function(df, th) {
  south |> 
    with(table(age_group, titer > th)) |> 
    as.data.frame() |> 
    pivot_wider(names_from = Var2, values_from = Freq) |> 
    mutate(tmp = map2(`TRUE`, `TRUE` + `FALSE`, binom_test2) |> 
             map(extract, c("estimate", "conf.int")) |>
             map(unlist) |> 
             map(setNames, c("p", "l", "u"))) |> 
    unnest_wider(tmp)
#}
```

```{r}
#south |> 
#  filter()
```


* model density() in 2D in order to see density of data points not only as a function 
of age but also as a function of titers

* doing clustering on the ANCOVA

```{r}
seroprevT1T2 <- function(x, th = 200, bs = "tp", k = 50, title = FALSE) {
  x |> 
    filter(time_point == "T1") |> 
    gam3((titer > th) ~ s(age, bs = bs, k = k), data = _) |> 
    predict3() |> 
    plot_pred(col = 2)
  
  x |> 
    filter(time_point == "T2") |>
    gam3((titer > th) ~ s(age, bs = bs, k = k), data = _) |> 
    predict3() |> 
    add_ribbon(3)
  
  if (title) x |> 
    pull(province) |> 
    first() |> 
    text(35, .2, labels = _)
}
```

```{r}
seroprev3timepoints <- function(x, th = 200, bs = "tp", k = 50, title = FALSE) {
  seroprevT1T2(x, th, bs, k, title)
  x |> 
    filter(time_point == "T3") |>
    gam3((titer > th) ~ s(age, bs = bs, k = k), data = _) |> 
    predict3() |> 
    add_ribbon(4)  
}
```

```{r fig-T1T2T3}
seroprev3timepoints(south)
```

```{r fig-T1T2}
seroprevT1T2(south)
```

```{r fig-seroprev_provinces_3timepoints, fig.width = 8.4, fig.height = 5 * 8.4 / 7}
opar <- par(mfrow = c(3,3), plt = c(.12, .97, .18, .95))
south |> 
  filter_out(province == "Kien Giang") |> 
  group_by(province) |> 
  group_split() |> 
  walk(seroprev3timepoints, title = TRUE)
par(opar)
```

```{r fig-seroprev_provinces, fig.width = 8.4, fig.height = 5 * 8.4 / 7}
opar <- par(mfrow = c(3,3), plt = c(.12, .97, .18, .95))
south |> 
  filter_out(province == "Kien Giang") |> 
  group_by(province) |> 
  group_split() |> 
  walk(plot_age_profile, 200)
par(opar)
```

* understanding the age profile of titers and seroprevalence. Maybe as an effect of
vaccination and waning immunity for vaccination

* Looking at 0-1 and 0-5 in order to see whether we see more detail signal as a
function of age

* test differences between sexes


